\texttt{mimix}: Reference-based imputation of missing data

### 2021-03-13 paper/paperRmimix.tex
Update paper after discussion with Kevin. Explore how wholly-missing observatins are handled in Stata and R

[ADD SOMETHING ON INDIVIDUALS WITH NO OBSERVED VALUES]

@@

Covariates are used to improve the plausibility of MAR assumptions and to improve the precision of imputed values \citep{ian:MItutorial}. 
In \texttt{mimix}, they are included in the multivariate normal model fitting as part of the variable $Y_0$. 
[KEVIN IS THIS CORRECT? MIGHT THEY BE WRONGLY CARRIED FORWARD UNDER LMCF?]

@@

\multicolumn{3}{l}{\em MCMC options} \\
& prior 	&	 Prior for the variance-covariance matrix when fitting multivariate normal distributions: Jeffreys (default), uniform or ridge.	\\
& burnin 	&	 Number of burn-in iterations when fitting multivariate normal distributions.	\\
& bbetween 	&	 Number of iterations between imputed data sets when fitting multivariate normal distributions.	\\
& mle 	&	 Logical option to use maximum likelihood parameter estimates instead of MCMC draw parameters. 
[NEED TO WRITE A COMMENT ABOUT THIS -- KEVIN, OR CAN WE UNDOCUMENT IT?]	\\ \hline
\end{tabular}

@@

The M imputed data sets are output concatenated as one large dataset appended to
the original unimputed dataset.
[KEVIN IS IT A DATA SET, A DATA FRAME, OR OTHER?]

@@

The data are missing after patients discontinue their randomised treatment. A treatment policy estimand is therefore estimated using reference-based imputation. For example, J2R is implemented as follows
\begin{verbatim}
asthmaJ2R <- mimix(data='asthma', covar=base, depvar=fev, treatvar=treat, idvar=id, timevar=time,
    method='J2R', reference=1, M=5, seed=101, prior=ridge, burnin=1000)
\end{verbatim}

[KEVIN, IT SEEMS COVAR MUST BE STRING BUT DEPVAR A VARNAME, IS THIS CORRECT? NEEDS A COMMENT?]

Note that the imputed data are in wide format. 
[KEVIN / ALL - I THINK THIS IS A PROBLEM!?]
The primary outcome can be analysed using an analysis of covariance by
\begin{verbatim}
library(mice)     
fit <- with(data= as.mids(asthmaJ2R), lm(fev.12 ~ treat + base))
summary(pool(fit))    
\end{verbatim}

[GIVE RESULTS]

@@

[TO DISCUSS -- TEXT OR TABLE? BOTH ARE INCLUDED BELOW.]

@@

Not using these options also means that the LMCF method can be compared with either ALMCF or OLMCF in the Five\_Macros: these are the same in the mimix setting with no covariate interacting with time.

@@

\begin{tabular}{p{.2\textwidth}p{.2\textwidth}p{.2\textwidth}p{.2\textwidth}}\hline
	&	Five macros in SAS	&	mimix in Stata	&	mimix in R	\\\hline
{\em Functionality}	&		&		&		\\
RBI	&	yes	&	yes	&	yes	\\
Causal model	&	yes	&	no	&	yes	\\
Delta method	&	yes	&	no	&	yes	\\\hline
{\em Options}	&		&		&		\\
Interim missing	&	MAR	&	RBI or MAR	&	MAR	\\
MCMC options	&	seed and thinning (=burnin)	&	seed, burnin and thinning	&	seed, burnin, thinning and prior	\\
Covariates interacting with treatment	&	never	&	always	&	always	\\
Covariates interacting with time	&	optional	&	never	&	never	\\
LMCF	&	several flavours: ALMCF, OLMCF, AFCMCF, OFCMCF	&	ALMCF only?	&	ALMCF only?	\\
No observed data	&	Warning	&	First mean is used	&	?	\\\hline
\end{tabular}

>>>>

\subsection{Individuals with no observed values}

[TEXT FOR DISCUSSION -- NEED TO WORK OUT HOW TO HANDLE THESE DISCREPANCIES]

Individuals with no observed values are easily imputed under MAR, but for the other methods they are currently a problem that we need to resolve.

We first consider J2R and CR. 
These methods should construct the imputation distribution wholly from the reference arm. 
They therefore correspond to MAR imputation if the wholly unobserved individual was in the reference arm. 
Both the Stata and R programs do this correctly for an individual in the reference arm, but for an individual in a non-reference arms, J2R and CR do not agree with each other.

We next consider CIR and LMCF. These are defined in terms of the last observed value, so they are formally undefined. 
The Stata program imputes CIR the same as J2R, while LMCF differs from all other methods, except that it agrees with MAR at the first visit.
The R program imputes CIR and J2R differently when in the non-reference arm.

@@

Covariates are used to improve the plausibility of MAR assumptions and to improve the precision of imputed values \citep{ian:MItutorial}. 
In \texttt{mimix}, they are included in the multivariate normal model fitting as part of the variable $Y_0$.

@@

\multicolumn{3}{l}{\em MCMC options} \\
& prior 	&	 Prior for the variance-covariance matrix when fitting multivariate normal distributions: Jeffreys (default), uniform or ridge.	\\
& burnin 	&	 Number of burn-in iterations when fitting multivariate normal distributions.	\\
& bbetween 	&	 Number of iterations between imputed data sets when fitting multivariate normal distributions.	\\
& mle 	&	 Logical option to use maximum likelihood parameter estimates instead of MCMC draw parameters. 
[WE'RE EXPLORING WHETHER WE CAN HIDE THIS OPTION, WHICH WAS ONLY WRITTEN TO PLEASE JONATHAN BARTLETT]	\\ \hline
\end{tabular}

@@

The M imputed data sets are output concatenated as one large data frame, comprising the 
original unimputed dataset concatenated with all the imptued data sets.
[CURRENTLY IN WIDE FORMAT, UNLIKE THE INPUT DATA AND UNLIKE STATA'S OUTPUT]

@@

The data are missing after patients discontinue their randomised treatment. 
A treatment policy estimand is therefore estimated using reference-based imputation, with the placebo arm as reference. 
For example, J2R is implemented as follows
\begin{verbatim}
asthmaJ2R <- mimix(data=asthma, covar='base', depvar=fev, treatvar=treat, idvar=id, 
    timevar=time, method='J2R', reference=1, M=5, seed=101, prior=ridge, burnin=1000)
\end{verbatim}

The primary outcome can be analysed using an analysis of covariance by
\begin{verbatim}
library(mice)     
fit <- with(data= as.mids(asthmaJ2R), lm(fev.12 ~ treat + base))
summary(pool(fit))    
\end{verbatim}
Results are as follows:
\begin{verbatim}
         term  estimate  std.error statistic       df      p.value
1 (Intercept) 0.5821281 0.21097369  2.759245 19.15041 1.242142e-02
2     treat.L 0.1585223 0.07119548  2.226578 81.64041 2.873017e-02
3        base 0.6843152 0.09180436  7.454060 28.90698 3.315976e-08
\end{verbatim}

@@

[TO DISCUSS -- TEXT, TABLE OR BOTH?]

@@

Not using these options also means that the LMCF method can be compared with either ALMCF or OLMCF in the Five\_Macros.
[I STILL HAVE TO WORK OUT WHICH WE HAVE IMPLEMENTED.]

@@

\begin{tabular}{p{.2\textwidth}p{.2\textwidth}p{.2\textwidth}p{.2\textwidth}}\hline
	&	Five macros in SAS	&	mimix in Stata	&	mimix in R	\\\hline
{\em Functionality}	&		&		&		\\
RBI	&	yes	&	yes	&	yes	\\
Causal model	&	yes	&	no	&	yes	\\
Delta method	&	yes	&	no	&	yes	\\\hline
{\em Options}	&		&		&		\\
Interim missing	&	MAR	&	RBI or MAR	&	MAR	\\
MCMC options	&	seed and thinning (=burnin)	&	seed, burnin and thinning	&	seed, burnin, thinning and prior	\\
Covariates interacting with treatment	&	never	&	always	&	always	\\
Covariates interacting with time	&	optional	&	always	&	always	\\
LMCF	&	several flavours: ALMCF, OLMCF, AFCMCF, OFCMCF	&	ALMCF only?	&	ALMCF only?	\\
No observed data	&	Warning	&	First mean is used	&	?	\\\hline
\end{tabular}
